---
title: "Main-usage-of-harmonizRLegacy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Main-usage-of-harmonizRLegacy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = T,
  results = "hide")

```


## Get started

The easiest way to get started with **harmonizRLegacy** is using this vignette 
as a demo with the current version on GitHub.

To download our package, you need to have an account on 
<a href="https://github.com/" target="_blank">GitHub</a> and create a 
<a href="https://github.com/settings/tokens" target="_blank">secure key</a> (in the 
format "ghp_xxx").

You can send us your Github name using the 
<a href="https://www.maelstrom-research.org/contact" target="_blank">following link</a> so 
we can add you to our authorized list of users.

Please note that all the demo files mentioned below are downloadable on an Opal demo environment.

```{r, eval=FALSE}

# To update the R package in your R environment you may first need to remove it, 
# and use the exit command quit() to terminate the current R session.
# Use the script below to install the R package.
devtools::install_github(
  repo = "https://github.com/maelstrom-research/harmonizRLegacy",
  auth_token = "ghp_xxx",
  force = TRUE)

library(harmonizRLegacy)

# additional packages needed in the demonstration
library(tidyverse)
library(opalr)

# please make sure that those libraries are installed on your R environment:
# 
# DT,         # tools,        # plotly,                    
# bookdown,   # utils,        # readr,                          
# dplyr,      # viridis,      # readxl,       
# fs,         # xlsx,         # rlang,        
# ggplot2,    # getPass,      # stringr,        
# grDevices,  # knitr,        # tibble,        
# janitor,    # car,          # tidyr,        
# opalr,      # lubridate     # tidytext,     

harmonizR_help()

```

## Overview

The aim of this vignette is to walk you through the harmonization process using 
the library **harmonizRLegacy** and illustrate its main functions.
**harmonizRLegacy** allows you to:

  * Create an R environment where you can place files in an organized structure
  * Identify the location and content of files in your environment
  * Generate bookdown reports of study-specific variables (input)
  * Read and interpret the data processing elements (harmonization potential and 
    data processing rules) and generate harmonized dataset(s) (output)
  * Generate bookdown reports of harmonized variables

## Set up your harmonization R environment

**Create an organized environment** to place your files correctly and control versions
of your project. A well-organized structure for your project will facilitate the 
harmonization work and avoid confusion between different versions of project files
(e.g., different versions of datasets, data dictionaries, DataSchema, and data processing
elements).

```{r, eval=FALSE}
# use case 1: create folder without versioning
harmo_env_create(project = "DEMO")

```

## Import the demonstration files

To test and use the library **harmonizRLegacy**, you can get access to demo 
files that are downloadable in our 
[Opal demo site](https://opal-demo.maelstrom-research.org/). 
Do not hesitate to contact us using the form previously mentioned above.


```{r, eval=FALSE}
o <- opal.login(username = 'administrator',
                password = 'password',
                url = 'https://opal-demo.maelstrom-research.org/')

opal_files_pull(o, from = "/home/administrator/DEMO/data_processing_elements", to = "DEMO")
opal_files_pull(o, from = "/home/administrator/DEMO/dataschema",               to = "DEMO")
opal_files_pull(o, from = "/home/administrator/DEMO/study_specific_datasets",  to = "DEMO")

```

## Index the files in your harmonization R environment

Create a [tibble](https://tibble.tidyverse.org/) listing all files in a folder 
(recursively) with some useful metadata. You can add a filter to your query to 
exclude folders from indexing. This tibble can also be used to quickly search for 
the location of a file. Finally, you can read the indexed files and generate objects 
in the R environment.

```{r, eval=FALSE}

# create index of files that are in a  folder
index_DEMO <- file_index_create(folder = "DEMO")

# search index for files matching a specific file path query
file_index_search(index = index_DEMO, file_path = "study_specific_datasets")

# read the files
file_index_read(index = index_DEMO, file_path = "study_specific_datasets")


```

## Generate visual report(s) of the study-specific variables

You can **generate quality assessment (QA) reports for study datasets**. To facilitate 
the assessment of study-specific datasets, you can generate html bookdown report 
showing summary statistics for each study-specific variable. The summary statistics 
and figures are generated according to their valueTypes. The variable valueTypes 
is automatically detected from the dataset and its data dictionary. You can 
specify a data dictionary, a grouping variable to get summary statistics by group, 
and the output folder (If not specified, the current folder is the default).

See [ggplot2](https://ggplot2.tidyverse.org/), [plotly](https://plotly.com/r/)
and [DT](https://rstudio.github.io/DT/) for further documentation about graphics and 
tables generated in the report.


```{r, eval=FALSE}

# use case 1: report of demo dataset TOKYO
study_visual_report(
  dataset = study_TOKYO,
  data_dict = dd_TOKYO_format_maelstrom_tagged,
  to = "DEMO/reports/TOKYO")

# use case 2: report of demo dataset TOKYO, grouped by gndr
study_visual_report(
  dataset = study_TOKYO,
  data_dict = dd_TOKYO_format_maelstrom_tagged,
  to = "DEMO/reports/TOKYO_gndr",group_by = "gndr",out = "ggplot2")

# re-index your files to include new files created
index_DEMO <- file_index_create(folder = "DEMO")

# read the bookdown
file_index_read(index_DEMO,file_path = "DEMO/reports/TOKYO_gndr/docs/index.html")

```

## Read and run the data processing rules and generate the harmonized dataset

**Read and run the data processing elements file** line by line or as a whole and
harmonize each study-specific dataset. Error messages are generated automatically
to help identify sources of errors in the data processing elements. 
Separate scripts for pre-processing datasets (before harmonization) or post-processing 
(after harmonization) can be integrated into the process.

```{r, eval=FALSE}

# Read the files
file_index_read(index_DEMO, file_path = "study_specific_datasets")
file_index_read(index_DEMO, file_path = "dataschema")
file_index_read(index_DEMO, file_path = "data_processing_elements")

# Example 1: run the process of harmonization with data processing elements containing errors
harmo_process(
  dataschema = DEMO_dataschema,
  data_proc_elem = `DEMO_data_processing_elements - with error`,
  output_harmo_folder = "DEMO/harmonized_datasets",
  name_report = "harmo_report")

# Example 2: use the function harmo_summary() to access the errors. 
# In your environment, you will find a report attached to harmonization process.
harmo_summary(harmo_report)
rm(harmo_report)

# After correction, re-run the process of harmonization with the corrected data 
# processing elements.
harmo_process(
  dataschema = DEMO_dataschema,
  data_proc_elem = `DEMO_data_processing_elements - final`,
  name_report = "final_report",
  output_harmo_folder = "DEMO/harmonized_datasets")

# re-index your files to include new files created
index_DEMO <- file_index_create(folder = "DEMO")

```

## Generate visual report of the harmonized variables

Generate **QA reports for the harmonized dataset(s)** to facilitate the assessment
of the harmonized dataset. A html bookdown report is generated showing summary statistics 
for each harmonized variable. The statistics generated for each variable are determined 
by the variable format, which is one of three types: numerical, text, or categorical.
Statistics are shown for the pooled harmonized dataset as well as by study. The DataSchema
must be specified. You can specify the output folder (if not specified, the working 
directory is the default).

```{r, eval=FALSE}

# pool your data in one tibble
harmonized_study_DEMO_table_DEMO <-
  harmonized_study_MELBOURNE_table_MELBOURNE %>%
  bind_rows(harmonized_study_PARIS_table_PARIS) %>%
  bind_rows(harmonized_study_TOKYO_table_TOKYO)

# use case 1: report of harmonized_study_DEMO_table_DEMO.
harmo_visual_report(
  harmonized_dataset = harmonized_study_DEMO_table_DEMO,
  dataschema = DEMO_dataschema,
  data_proc_elem = `DEMO_data_processing_elements - work in progress`,
  to = "DEMO/reports/harmonized_DEMO"
  # group_by = "adm_study"
)

# re-index your files to include new files created
index_DEMO <- file_index_create(folder = "DEMO")

# read the bookdown (or open the file named 'index.html' in your browser)
file_index_read(index_DEMO,file_path = "DEMO/reports/harmonized_DEMO/docs/index.html")

```
